---
output: html_document
---


```{r setup, include=FALSE}
library(glptools)

glp_load_packages()

library(lubridate)
library(sf)
library(leaflet)
library(viridis)
```

# Create data set

## Read SNAP data

Outputs:
-snap_establishments: a data frame of SNAP retailers in Louisville and peer cities
-snap_establishments_df: a data frame of large retailers that are used to calculate food deserts. This includes retailers slightly outside of counties as well.

```{r read_snap}
snap_establishments <- read_csv("Historical SNAP Retailer Locator Data as of 20211231.csv")

# Clean data frame and convert opening and closing to dates.
# We have location data for all stores that existed or opened starting in 2006.
# We are missing location data for most older stores, so remove them from the data.

snap_establishments %<>%
  transmute(
    ID = `Record ID`,
    name = `Store Name`,
    type = `Store Type`,
    street = `Street Name`,
    city = City,
    county = County,
    state_abbr = State,
    zip = `Zip Code`,
    lat = Latitude,
    lon = Longitude,
    open = mdy(`Authorization Date`),
    close = mdy(`End Date`)) %>%
  filter(is.na(close) | year(close) >= 2006)

# Convert the data frame to a spatial object
snap_establishments %<>% st_as_sf(coords = c("lon", "lat"), crs = 4326)

# Since people can access food in other counties, let's create a "nearby" dataset to include food options within a mile of county borders.
FIPS_map <- map_county_peers %>%
  pull_peers(add_info = T) %>%
  filter(current == 1) %>%
  st_buffer(dist = 1609)

FIPS_map_lou <- map_county_peers %>%
  filter(FIPS == "21111") %>%
  st_buffer(dist = 1609 * 5)

# Chop down data to near counties
snap_buffer = snap_establishments[FIPS_map,]
snap_buffer_lou = snap_establishments[FIPS_map_lou,]

# Add FIPS column to data frame. Format St. Louis city and county to match the FIPS_info first. Keep the data only for current peers.
snap_establishments %<>%
  mutate(
    county = case_when(
      county == "ST LOUIS" ~ "St. Louis County",
      county == "ST LOUIS CITY" ~ "St. Louis City",
      TRUE ~ str_to_title(county))) %>%
  select(-city) %>%
  left_join(FIPS_info, by = c("state_abbr", "county")) %>%
  filter(!is.na(FIPS),
         current == 1) %>%
  select(ID, name, type, open, close, FIPS, geometry)

# For the purposes of food deserts, we want to focus on large groceries, supermarkets, and superstores.
snap_establishments_fd <- snap_establishments %>%
  filter(type %in% c("Large Grocery Store", "Supermarket", "Super Store"))


# Clean up the outside-of-county food options and add them to the food desert dataset.
snap_nearby <- snap_buffer %>%
  anti_join(st_drop_geometry(snap_establishments), by = c("ID", "open")) %>%
  filter(type %in% c("Large Grocery Store", "Supermarket", "Super Store")) %>%
  select(ID, name, type, open, close, geometry)

snap_nearby$FIPS <- FIPS_map$FIPS[st_nearest_feature(snap_nearby, FIPS_map)]

snap_nearby_lou <- snap_buffer_lou %>%
  anti_join(st_drop_geometry(snap_establishments), by = c("ID", "open")) %>%
  filter(type %in% c("Large Grocery Store", "Supermarket", "Super Store")) %>%
  select(ID, name, type, open, close, geometry)

snap_nearby_lou$FIPS <- "21111"

snap_establishments_fd %<>%
  bind_rows(snap_nearby)

# Create Louisville-specific data set to run through hereR multiple times.
snap_lou <- snap_establishments_fd %>%
  filter(FIPS == "21111")

snap_nearby_lou <- snap_nearby %>% filter(FIPS == "21111")

snap_lou_all <- bind_rows(snap_lou, snap_nearby_lou)

rm(FIPS_map, FIPS_map_lou, snap_nearby, snap_buffer, snap_buffer_lou)

```

# Geocode and save data

Geocoding is done using the hereR package to generate isolines that show who can access grocery stores within a certain distance of their home.

The peer data set is analyzed using a standard 1-mile radius, but we generate additional isolines for Louisville to show grocery stores within a 1/2 mile, 2 miles, 3 miles, 4 miles, and 5 miles.

```{r geocode_lou, eval=FALSE}


library(hereR)
set_key("WoYNTJckWFwljeKbygY_G1L1uZ46k-DMJYWtlloRIXg")


grocery_location <- data.frame(
  ID = 1,
  name = "Louisville Community Grocery",
  type = "Large Grocery Store",
  open = "2022-01-08",
  close = NA_character_,
  FIPS = "21111",
  lat = "38.243894",
  lon = "-85.747014") %>%
  st_as_sf(coords = c("lon", "lat"), crs = 4326)

# Community grocery
lou_grocery <- 
  isoline(grocery_location,
          arrival = FALSE,
          range = 1609,
          range_type = "distance",
          routing_mode = "short",
          traffic = FALSE,
          optimize = "quality",
          aggregate = F,
          url_only = F)

save(lou_grocery, file = "community_grocery_isoline.RData")


driving_dist_half <- isoline(snap_lou,
                             arrival = FALSE,
                             range = 805,
                             range_type = "distance",
                             routing_mode = "short",
                             traffic = FALSE,
                             optimize = "quality",
                             aggregate = F,
                             url_only = F)

lou_half <- driving_dist_half %>%
  transmute(
    ID = snap_lou$ID,
    open = snap_lou$open,
    geometry)

save(lou_half, file = "louisville_food_isolines_half_clean.RData")


driving_dist_1    <- isoline(snap_lou,
                             arrival = FALSE,
                             range = 1609,
                             range_type = "distance",
                             routing_mode = "short",
                             traffic = FALSE,
                             optimize = "quality",
                             aggregate = F,
                             url_only = F)

save(driving_dist_1, file = "louisville_food_isolines_1.RData")

lou_1 <- driving_dist_1 %>%
  transmute(
    ID = snap_lou$ID,
    open = snap_lou$open,
    geometry)

save(lou_1, file = "louisville_food_isolines_1_clean.RData")

driving_dist_2    <- isoline(snap_lou,
                             arrival = FALSE,
                             range = 1609 * 2,
                             range_type = "distance",
                             routing_mode = "short",
                             traffic = FALSE,
                             optimize = "quality",
                             aggregate = F,
                             url_only = F)

lou_2 <- driving_dist_2 %>%
  transmute(
    ID = snap_lou$ID,
    open = snap_lou$open,
    geometry)

save(lou_2, file = "louisville_food_isolines_2_clean.RData")


driving_dist_3    <- isoline(snap_lou,
                             arrival = FALSE,
                             range = 1609 * 3,
                             range_type = "distance",
                             routing_mode = "short",
                             traffic = FALSE,
                             optimize = "quality",
                             aggregate = F,
                             url_only = F)

lou_3 <- driving_dist_3 %>%
  transmute(
    ID = snap_lou$ID,
    open = snap_lou$open,
    geometry)

save(lou_3, file = "louisville_food_isolines_3_clean.RData")

driving_dist_4    <- isoline(snap_lou,
                             arrival = FALSE,
                             range = 1609 * 4,
                             range_type = "distance",
                             routing_mode = "short",
                             traffic = FALSE,
                             optimize = "quality",
                             aggregate = F,
                             url_only = F)

lou_4 <- driving_dist_4 %>%
  transmute(
    ID = snap_lou$ID,
    open = snap_lou$open,
    geometry)

save(lou_4, file = "louisville_food_isolines_4_clean.RData")


driving_dist_5    <- isoline(snap_lou,
                             arrival = FALSE,
                             range = 1609 * 5,
                             range_type = "distance",
                             routing_mode = "short",
                             traffic = FALSE,
                             optimize = "quality",
                             aggregate = F,
                             url_only = F)

lou_5 <- driving_dist_5 %>%
  transmute(
    ID = snap_lou$ID,
    open = snap_lou$open,
    geometry)

save(lou_5, file = "louisville_food_isolines_5_clean.RData")

save(lou_half, lou_1, lou_2, lou_3, lou_4, lou_5,
     file = "louisville_food_isolines.RData")

snap_establishments_fd_peer <- snap_establishments_fd %>% filter(FIPS != "21111")

start = 1
end = 100
  
for(i in 1:30) {
  
  file_name = "peer_isolines_" %p% i %p% ".RData"
  
  print(i) 
  print(start)
  print(end)
  print(file_name)
  
  temp    <- isoline(snap_establishments_fd_peer[start:end,],
                     arrival = FALSE,
                     range = 1609,
                     range_type = "distance",
                     routing_mode = "short",
                     traffic = FALSE,
                     optimize = "quality",
                     aggregate = F,
                     url_only = F)
  
  save(temp, file = file_name)
  
  start = start + 100
  end = end + 100
  
  if(end >2900) end <- 2947
}

```

## Load and clean data

SNAP data - 

Louisville data:
--lou_half, lou_1, lou_2, lou_3, lou_4, lou_5
--peer isolines

Map data:



```{r, eval=FALSE}

# Load Louisville data

load("louisville_food_isolines_1_clean2.RData")
load("community_grocery_isoline.RData")

lou_1 %<>%
  left_join(st_drop_geometry(snap_establishments), by = c("ID", "open")) %>%
  filter(!is.na(name))
  

file_names <- paste0("peer_isolines_", 1:30, ".RData")

fxn <- function(file_name) {
  load(file_name)
  
  temp
}

peer_isolines <- map_dfr(file_names, fxn)

# Add grocery

peer_isolines2 <- peer_isolines %>% bind_rows(lou_grocery)

peer_snap_df <- snap_establishments_fd %>% 
  filter(FIPS != "21111") %>%
  st_drop_geometry()

peer_isolines %<>%
  bind_cols(peer_snap_df)

FIPS_map <- glptools::map_county_peers %>%
  pull_peers(add_info = T) %>%
  filter(current == 1) %>%
  group_by(city) %>%
  summarize() %>%
  left_join(FIPS_df_one_stl, by = "city") %>%
  select(FIPS, geometry)

lou_map <- FIPS_map %>%
  filter(FIPS == "21111")


peer_isolines %<>%
  select(ID, open, name, type, close, FIPS, geometry) %>% 
  bind_rows(lou_1)

lou_grocery %<>%
  mutate(open = lubridate::ymd("2020-01-01"), close = NA_Date_, FIPS = "21111", 
         type = "Large Grocery Store")

peer_isolines %<>% bind_rows(lou_grocery)


library(rmapshaper)
t <- ms_simplify(peer_isolines, keep = 0.05, keep_shapes = FALSE)

t2=t %>% st_make_valid()

t3=t2 %>%
  #filter(is.na(close)) %>%
  mutate(FIPS = if_else(FIPS %in% c("29189", "29510"), "MERGED", FIPS))

t3 = sf::st_transform(t3, crs = 3857)
t3 = sf::st_transform(t3, crs = 4326)

t3 %<>% st_make_valid()
  
results_skeleton <- crossing(
  FIPS = FIPS_map$FIPS,
  year = 2011:2021)

rm(final_df)

for(f in 1:nrow(results_skeleton)) {
  
  print(f)
  
  this_FIPS <- results_skeleton[f, "FIPS"] %>% pull(FIPS)
  this_year <- results_skeleton[f, "year"] %>% pull(year)
  
  t4 <- t3 %>%
    filter(year(open) <= this_year,
           year(close) > this_year | is.na(close),
           FIPS == this_FIPS)
  
  # Merge FD info
  t5=t4 %>%
    st_union()
  
  t6 = t5 %>%
    st_make_valid()
  
  # Get tract map
  
  tracts <- glptools::map_tract_all_10

  tracts <- tracts %>% 
    mutate(FIPS = if_else(FIPS %in% c("29189", "29510"), "MERGED", FIPS)) %>%
    filter(FIPS == this_FIPS)
  
  # Intersect
  
  tract_df <- st_intersection(tracts, t6)
  
  tract_full_area <- tracts %>%
    mutate(area = st_area(.)) %>%
    st_drop_geometry()
  
  tract_df %<>%
    mutate(non_fd_area = st_area(.)) %>%
    st_drop_geometry()

  output_df <- tract_full_area %>%
    left_join(tract_df, by = c("FIPS", "tract")) %>%
    transmute(
      tract,
      year = this_year,
      pct_food_desert = as.numeric((area - non_fd_area) / area) %>%
        replace_na(1))
  
  final_df <- assign_row_join(final_df, output_df)
  
}

save(final_df, file = "final_df.RData")

```

# Graphs 

## Louisville widget

```{r, eval=FALSE}
library(shiny)
library(shinyWidgets)

starting_data <- lou_1 %>%
  left_join(st_drop_geometry(snap_establishments_fd), by = c("ID", "open")) %>%
  filter(is.na(close), !is.na(type))

pal <- leaflet::colorFactor(
  palette = viridis(3),
  domain = c("Super Store", "Supermarket", "Large Grocery Store"))

starting_map <- leaflet(starting_data) %>%
  addTiles() %>%
  addPolygons(data = filter(map_county, FIPS == "21111"),
            color = "black",
            opacity = 1,
            weight = 2,
            fillOpacity = 0) %>%
  addPolygons(
    data = starting_data,
    layerId = "fd_1",
    color = ~pal(type),
               opacity = 0.7,
               weight = 2,                   
               label = ~name,
               labelOptions = labelOptions(style = 
                list("font-weight" = "normal", 
                     "font-family" = "Montserrat", 
                     padding = "3px 8px"),
                textsize = "15px",
                direction = "auto")) %>%
 addLegend(title = "Store Type", labels = c("Super Store", "Supermarket", "Large Grocery Store"), colors = viridis(3))

# Define UI for application that draws a histogram
ui <- flowLayout(
  
    # Sidebar with a slider input for number of bins
  sliderTextInput(
    inputId = "miles",
    label = "Miles:", 
    selected = 1,
    grid = TRUE,
    choices = c(0.5, 1, 2, 3, 4, 5)),
      
    # Show a plot of the generated distribution
   leafletOutput("fd_map", "400px")
)

# Define server logic required to draw a histogram
server <- function(input, output) {

  fd_dataset <- reactive({
    
    if (input$miles == 0.5) this_fd <- lou_half
    if (input$miles == 1) this_fd <- lou_1
    if (input$miles == 2) this_fd <- lou_2
    if (input$miles == 3) this_fd <- lou_3
    if (input$miles == 4) this_fd <- lou_4
    if (input$miles == 5) this_fd <- lou_5
    
   this_fd %<>%
    left_join(st_drop_geometry(snap_establishments_fd), by = c("ID", "open")) %>%
    filter(is.na(close), !is.na(type))
    
    this_fd
  })
  
  output$fd_map <- renderLeaflet(starting_map)
  
  observeEvent(fd_dataset(), {
    leafletProxy("fd_map") %>%
      clearShapes() %>%
      
      # add shape color and white lines for all polygons
      addPolygons(
        data = fd_dataset(),
        color = ~pal(type),
               opacity = 0.7,
               weight = 2,                   
               label = ~name,
               labelOptions = labelOptions(style = 
                list("font-weight" = "normal", 
                     "font-family" = "Montserrat", 
                     padding = "3px 8px"),
                textsize = "15px",
                direction = "auto"))

  })
}

# Run the application
shinyApp(ui = ui, server = server)



# knitr::include_app("https://yihui.shinyapps.io/miniUI/",
#   height = "600px")

starting_map

```

# Other data

## Vehicle

```{r}
vehicle_variables   <- build_census_var_df("acs5", "B08201")

vehicles_available <- get_census(vehicle_variables, "tract_all")

vehicles_available %<>%
  mutate(no_vehicle = case_when(
    str_detect(variable, "B08201_002") ~ TRUE,
    str_detect(variable, "_003|_004|_005|_006") ~ FALSE,
    TRUE ~ NA))

vehicles_map <- process_census(vehicles_available, cat_var = "no_vehicle", output_name = "no_vehicles_available")

vehicles_map_20 <- vehicles_map %>%
  filter(year == 2018) %>%
  pivot_vartype_wider(no_vehicles_available) %>%
  left_join(rename(tract20_tract10, pct_tract = percent), by = c("tract" = "tract20")) %>%
  group_by(tract10, year, variable) %>%
  summarize(
    estimate = sum(estimate * pct_tract), 
    population = sum(population * pct_tract),
    percent = estimate / population * 100,
    .groups = "drop") %>%
  pivot_vartype_longer() %>%
  transmute(
    tract = tract10,
    year,
    sex = "total",
    race = "total",
    var_type,
    no_vehicles_available)

vehicles_map %<>%
  filter(year != 2018) %>%
  bind_rows(vehicles_map_20)
```

## Bus stops

```{r}
bus_stops <- sf::read_sf("Louisville_Metro_Area_KY__TARC_Bus_Stops")
bus_routes <- sf::read_sf("Louisville_Metro_Area_KY_TARC_Bus_Routes_")

#library(hereR)
#set_key("-DMJYWtlloRIXg")

# start = 1
# end = 100
#   
# for(i in 1:39) {
#   
#   file_name = "bus_stop_isolines/bus_" %p% i %p% ".RData"
#   
#   print(i) 
#   print(start)
#   print(end)
#   print(file_name)
#   
#   temp    <- isoline(bus_stops[start:end,],
#                      arrival = FALSE,
#                      range = 400,
#                      range_type = "distance",
#                      routing_mode = "short",
#                      transport_mode = "pedestrian",
#                      traffic = FALSE,
#                      optimize = "quality",
#                      aggregate = F,
#                      url_only = F)
#   
#   save(temp, file = file_name)
#   
#   start = start + 100
#   end = end + 100
#   
#   if(end > 3876) end <- 3876
# }
# 
bus_files <- list.files("bus_stop_isolines")


for(i in 1:length(bus_files)) {
  load("bus_stop_isolines/" %p% bus_files[i])

  bus_stop_isolines <- assign_row_join(bus_stop_isolines, temp)

}

save(bus_stop_isolines, file = "bus_stop_isolines.RData")

# 
# 
# start = 1
# end = 100
# 
# output_list <- list()
# 
# for(i in 1:39) {
# 
#   print(i)
#   print(start)
#   print(end)
# 
#   temp_union <- bus_stop_isolines[start:end,]
# 
#   temp_union %<>%
#     st_union()
# 
#   if(i %not_in% c(15, 16, 21, 29, 31)) {
# 
# 
#     temp_union %<>%
#       rmapshaper::ms_simplify()
#   }
# 
# 
#   output_list <- c(output_list, temp_union)
# 
#   start = start + 100
#   end = end + 100
#   if(end > 3876) end <- 3876
# 
# }

# 
# t = flatten(output_list) %>% flatten() %>%
#   map
# 
# t2=st_as_sf(as.data.frame(t[[1]]), coords = c("V1", "V2"), crs = 4326)
# 
#   
# t %>% map(st_as_sf(meuse, coords = c("x", "y"), crs = 28992, agr = "constant"))
#             
#   
#   st_as_sf(coords = c("long", "lat"), crs = 4326)
# 
# 
# for(i in 1:length(output_list)) {
#   
#   
# }
# 
# 
# t=list(output_list, makeUniqueIDs = T) %>% 
#   flatten() %>% 
#   do.call(rbind, .)
# 
# data.frame(geometry = output_list)
# 
# st_as_sf(t)

# bus_stop_isolines_union <- bus_stop_isolines %>%
#   st_union()
# 
# bus_stop_isolines %<>%
#   summarize()

```

## Poverty

```{r}
poverty_ratio_vars <- build_census_var_df("acs5", "C17002")

poverty_ratio <- get_census(poverty_ratio_vars, "tract_all")

poverty_ratio %<>%
  mutate(pov_ratio = case_when(
    str_detect(variable, "_002") ~ "p50",
    str_detect(variable, "_003") ~ "p100",
    str_detect(variable, "_004") ~ "p125",
    str_detect(variable, "_005") ~ "p150",
    str_detect(variable, "_006") ~ "p185",
    str_detect(variable, "_007") ~ "p200",
    str_detect(variable, "_008") ~ "p200_plus",
    TRUE ~ NA_character_))

poverty_map <- process_census(poverty_ratio, cat_var = "pov_ratio", output_name = "poverty_ratio")

poverty_map_20 <- poverty_map %>%
  filter(year == 2018) %>%
  pivot_vartype_wider(p100:p50) %>%
  left_join(rename(tract20_tract10, pct_tract = percent), by = c("tract" = "tract20")) %>%
  group_by(tract10, year, variable) %>%
  summarize(
    estimate = sum(estimate * pct_tract), 
    population = sum(population * pct_tract),
    percent = estimate / population * 100,
    .groups = "drop") %>%
  pivot_vartype_longer() %>%
  transmute(
    tract = tract10,
    year,
    sex = "total",
    race = "total",
    var_type,
    p50,
    p100,
    p125,
    p150,
    p185,
    p200,
    p200_plus)

poverty_map %<>%
  filter(year != 2018) %>%
  bind_rows(poverty_map_20)
```

## SNAP

```{r}
snap_vars <- build_census_var_df("acs5", "B22001")

snap_df <- get_census(snap_vars, "tract_all")

snap_df %<>%
  mutate(snap_hh = case_when(
    str_detect(variable, "_002") ~ TRUE,
    str_detect(variable, "_005") ~ FALSE,
    TRUE ~ NA))

snap_map <- process_census(snap_df, cat_var = "snap_hh", output_name = "snap")

snap_map_20 <- snap_map %>%
  filter(year == 2018) %>%
  pivot_vartype_wider(snap) %>%
  left_join(rename(tract20_tract10, pct_tract = percent), by = c("tract" = "tract20")) %>%
  group_by(tract10, year, variable) %>%
  summarize(
    estimate = sum(estimate * pct_tract), 
    population = sum(population * pct_tract),
    percent = estimate / population * 100,
    .groups = "drop") %>%
  pivot_vartype_longer() %>%
  transmute(
    tract = tract10,
    year,
    sex = "total",
    race = "total",
    var_type,
    snap)

snap_map %<>%
  filter(year != 2018) %>%
  bind_rows(snap_map_20)
```

## Age

```{r}
age_vars <- build_census_var_df("acs5", "B01001")

age_vars %<>%
  filter(race == "total", age_group != "all")

age_df <- get_census(age_vars, "tract_all")

# Create a categorical variable for seniors and replace age groups with "all" to work with process census
age_df %<>%
  mutate(
    senior = case_when(
      age_low >= 65 ~ TRUE,
      age_low < 65 ~ FALSE,
      TRUE ~ NA),
    age_group = "all") 

age_map <- process_census(age_df, cat_var = "senior", output_name = "senior2")

age_map %<>%
  rename(senior = senior2)

age_map_20 <- age_map %>%
  filter(year == 2018) %>%
  pivot_vartype_wider(senior) %>%
  left_join(rename(tract20_tract10, pct_tract = percent), by = c("tract" = "tract20")) %>%
  group_by(tract10, year, variable) %>%
  summarize(
    estimate = sum(estimate * pct_tract), 
    population = sum(population * pct_tract),
    percent = estimate / population * 100,
    .groups = "drop") %>%
  pivot_vartype_longer() %>%
  transmute(
    tract = tract10,
    year,
    sex = "total",
    race = "total",
    var_type,
    senior)

age_map %<>%
  filter(year != 2018) %>%
  bind_rows(age_map_20)
```


## Combine data

```{r, eval=FALSE}
save(vehicles_map, poverty_map, snap_map, age_map, file = "census_exports.RData")

load("census_exports.RData")

load("louisville_food_isolines.RData")
load("food_desert_1mi.RData")
load("final_df.RData")


# Vehicles
vehicle_fd <- vehicles_map %>%
  mutate(FIPS = str_sub(tract, 1, 5)) %>%
  filter(var_type %in% c("estimate", "population")) %>%
  pivot_vartype_wider(no_vehicles_available) %>%
  mutate(year = year + 2)

vehicle_fd %<>%
  filter(year %in% 2010:2020)

vehicle_fd %<>%
  left_join(final_df, by = c("tract", "year")) %>%
  mutate(fd_pop = estimate * pct_food_desert) 

vehicle_fd_county <- vehicle_fd %>%
  group_by(FIPS, year) %>%
  summarize(
    food_desert = sum(fd_pop) / sum(population, na.rm = T) * 100, 
    non_food_desert = sum(estimate - fd_pop) / sum(population, na.rm = T) * 100, 
    pct_food_insecure = food_desert / (food_desert + non_food_desert) * 100, .groups = "drop") %>%
  stl_merge(food_desert:pct_food_insecure, simple = T)
  
vehicle_fd_map  <- vehicle_fd %>%
  transmute(tract, year, 
            food_desert = fd_pop / population * 100,
            non_food_desert = (estimate - fd_pop) / population * 100, 
            pct_food_insecure = food_desert / (food_desert + non_food_desert) * 100)

# Low income

poverty_fd_est <- poverty_map %>%
  mutate(FIPS = str_sub(tract, 1, 5)) %>%
  filter(var_type %in% "estimate") %>%
  transmute(FIPS, tract, year, var_type, low_income = p50+p100+p125+p150+p185+p200) %>%
  mutate(year = year + 2)

poverty_fd_pop <- poverty_map %>%
  mutate(FIPS = str_sub(tract, 1, 5)) %>%
  filter(var_type %in% "population") %>%
  transmute(FIPS, tract, year, var_type, low_income = p50) %>%
  mutate(year = year + 2)

poverty_fd <- bind_rows(poverty_fd_est, poverty_fd_pop)

poverty_fd %<>%
  filter(year %in% 2010:2020)

poverty_fd %<>%
  pivot_vartype_wider(low_income) %>%
  left_join(final_df, by = c("tract", "year")) %>%
  mutate(fd_pop = estimate * pct_food_desert) 
  
poverty_fd_county <- poverty_fd %>%
  group_by(FIPS, year) %>%
  summarize(
    food_desert = sum(fd_pop) / sum(population, na.rm = T) * 100, 
    non_food_desert = sum(estimate - fd_pop) / sum(population, na.rm = T) * 100, 
    pct_food_insecure = food_desert / (food_desert + non_food_desert) * 100, .groups = "drop") %>%
  stl_merge(food_desert:pct_food_insecure, simple = T)
  
poverty_fd_map  <- poverty_fd %>%
  transmute(tract, year, 
            food_desert = fd_pop / population * 100,
            non_food_desert = (estimate - fd_pop) / population * 100, 
            pct_food_insecure = food_desert / (food_desert + non_food_desert) * 100)

# SNAP

snap_fd <- snap_map %>%
  mutate(FIPS = str_sub(tract, 1, 5)) %>%
  filter(var_type %in% c("estimate", "population")) %>%
  pivot_vartype_wider(snap) %>%
  mutate(year = year + 2)

snap_fd %<>%
  filter(year %in% 2010:2020)

snap_fd %<>%
  left_join(final_df, by = c("tract", "year")) %>%
  mutate(fd_pop = estimate * pct_food_desert) 

snap_fd_county <- snap_fd %>%
  group_by(FIPS, year) %>%
  summarize(
    food_desert = sum(fd_pop) / sum(population, na.rm = T) * 100, 
    non_food_desert = sum(estimate - fd_pop) / sum(population, na.rm = T) * 100, 
    pct_food_insecure = food_desert / (food_desert + non_food_desert) * 100, .groups = "drop") %>%
  stl_merge(food_desert:pct_food_insecure, simple = T)
  
snap_fd_map  <- snap_fd %>%
  transmute(tract, year, 
            food_desert = fd_pop / population * 100,
            non_food_desert = (estimate - fd_pop) / population * 100, 
            pct_food_insecure = food_desert / (food_desert + non_food_desert) * 100)


# Senior
age_fd <- age_map %>%
  mutate(FIPS = str_sub(tract, 1, 5)) %>%
  filter(var_type %in% c("estimate", "population")) %>%
  pivot_vartype_wider(senior) %>%
  mutate(year = year + 2) %>%
  filter(sex == "total")

age_fd %<>%
  filter(year %in% 2010:2020)

age_fd %<>%
  left_join(final_df, by = c("tract", "year")) %>%
  mutate(fd_pop = estimate * pct_food_desert) 

age_fd_county <- age_fd %>%
  group_by(FIPS, year) %>%
  summarize(
    food_desert = sum(fd_pop) / sum(population, na.rm = T) * 100, 
    non_food_desert = sum(estimate - fd_pop) / sum(population, na.rm = T) * 100, 
    pct_food_insecure = food_desert / (food_desert + non_food_desert) * 100, .groups = "drop") %>%
  stl_merge(food_desert:pct_food_insecure, simple = T)
  
age_fd_map  <- age_fd %>%
  transmute(tract, year, 
            food_desert = fd_pop / population * 100,
            non_food_desert = (estimate - fd_pop) / population * 100, 
            pct_food_insecure = food_desert / (food_desert + non_food_desert) * 100)

all_fd <- age_map %>%
  mutate(FIPS = str_sub(tract, 1, 5)) %>%
  filter(FIPS == "21111", var_type == "population", year == 2018) %>%
  left_join(final_df, by = c("tract", "year")) %>%
  summarize(
    percent_fd = sum(senior * pct_food_desert, na.rm = T) / 
      sum(senior, na.rm = T))

save(poverty_fd_county, snap_fd_county, vehicle_fd_county, age_fd_county, poverty_fd_map, snap_fd_map, vehicle_fd_map, age_fd_map, file= "food desert exports.RData")
```




```{r, eval=FALSE}

glp_load_packages(graphs = T)
showtext_auto()
font_add("Museo Sans", "../../../../MuseoSans_300.otf")
font_add("Museo Sans 300 Italic", "../../../../MuseoSans_300_Italic.otf")
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, 
                      dev.args=list(bg="transparent"), fig.width=30, fig.height=24)

png("Vehicle Desert.png", 3000, 2000, res = 200)
glptools::trend(
  vehicle_fd_county,
  food_desert,
  plot_title = "Low Access and No Vehicle",
  y_title = "Percent")
dev.off()

png("Poverty Desert.png", 3000, 2000, res = 200)
glptools::trend(
  poverty_fd_county,
  food_desert,
  plot_title = "Low Access and Poverty",
  y_title = "Percent")
dev.off()

png("SNAP Desert.png", 3000, 2000, res = 200)
glptools::trend(
  snap_fd_county,
  food_desert,
  plot_title = "Low Access and SNAP",
  y_title = "Percent")
dev.off()

png("Age Desert.png", 3000, 2000, res = 200)
glptools::trend(
  age_fd_county,
  food_desert,
  plot_title = "Senior and SNAP",
  y_title = "Percent")
dev.off()

glptools::ranking(
  snap_fd_county,
  order = "Ascending",
  food_desert,
  plot_title = "Vehicle Desert",
  y_title = "Percent")


df <- glptools::ranking_data(
  poverty_fd_county,
  food_desert) %>%
  pull_peers(add_info = T)

write_csv(df, "temp.csv")

save(poverty_fd_county, snap_fd_county, vehicle_fd_county, age_fd_county, poverty_fd_map, snap_fd_map, vehicle_fd_map, age_fd_map, file= "food desert exports.RData")

load("food desert exports.RData")

```

```{r, eval=FALSE}
vehicle_map_19 <- vehicles_map %>%
  filter(var_type == "estimate", year == 2017)

poverty_map_19 <- poverty_map %>%
  filter(var_type == "estimate", year == 2017) 
  

jc_union = map_tract %>%
  st_union()

grocery_union = lou_1 %>%
  st_union() %>%
  st_make_valid()

grocery_map = map_tract %>%
  st_intersection(grocery_union)

t1=grocery_map %>%
  mutate(area_grocery = st_area(.))

t2=map_tract %>%
  mutate(area_total = st_area(.))


t2 %<>%
  st_drop_geometry() %>%
  left_join(st_drop_geometry(t1)) %>%
  mutate(covered = 1-as.numeric(area_grocery / area_total) %>%
           replace_na(0)) %>%
  select(tract, covered)

poverty_map_19 %<>%
  left_join(t2, by = "tract")


pov_sum <- poverty_map_19 %>%
  summarize(across(p100:p50, sum))

pov_sum_fd <- poverty_map_19 %>%
  summarize(across(p100:p50, ~sum(. * covered)))



leaflet(grocery_map) %>%
  addPolygons()
```

# Final maps

```{r, eval=FALSE}
load("louisville_food_isolines.RData")
load("food_desert_1mi.RData")

this_data <- lou_1 %>%
  left_join(st_drop_geometry(snap_establishments), by = c("ID", "open")) %>%
  filter(is.na(close))

library(smoothr)

pal <- leaflet::colorFactor(
  palette = viridis(3),
  domain = c("Large Grocery Store", "Super Store", "Supermarket"))


lou_food_map <- this_data

leaflet(this_data) %>%
  addTiles("https://server.arcgisonline.com/ArcGIS/rest/services/Canvas/World_Light_Gray_Base/MapServer/tile/{z}/{y}/{x}.png") %>%
  addPolygons(color = ~pal(type),
                   opacity = 0.7,
                   weight = 2,                   
                   label = ~name,
                   labelOptions = labelOptions(style = 
                    list("font-weight" = "normal", 
                         "font-family" = "Montserrat", 
                         padding = "3px 8px"),
                    textsize = "15px",
                    direction = "auto")) %>%
    # addPolygons(data = filter(glptools::map_msa_lou, FIPS != "21111"),
    #           color = "white",
    #           opacity = 1,
    #           weight = 0,
    #           fillOpacity = 1) %>%
  
  addPolygons(data = filter(map_county, FIPS == "21111"),
              color = "black",
              opacity = 1,
              weight = 2,
              fillOpacity = 0) %>%
 addLegend(title = "Store Type", labels = c("Large Grocery Store", "Super Store", "Supermarket"), colors = viridis(3))

save(lou_food_map, file = "Louisville map.RData")


```
